$p$-adic precision

### 2014-02-13 padicprec.tex
Generalization of main Lemma to manifolds

We introduce the following definition:

@@

\subsection{The case of locally analytic functions}

@@

Let us assume now that $f$ is locally analytic on $U$ and that $x$ and 
$H$ are chosen so that Lemma \ref{lem:main} applies. Then $f(x+H) = f(x) 
+ f'(x)(H)$ and the smallest possible $H'$ is $f'(x)(H)$. In other 
words, the function {\tt f} generates artificial loss of precision if 
the precision $H'$ it outputs is strictly larger that $f'(x)(H)$.
The aim of this section is to explain how, under the above hypothesis, 
one can implement the function {\tt f} so that it always outputs the
optimal precision $O(H')$ with $H' = f'(x)(H)$.

@@

There is nevertheless an important issue with this approach: in order to 
be sure that Lemma \ref{lem:main} applies, we need \emph{a priori} to 
compute the exact values of all $x_i$'s, which is of course not possible! 
Assuming that $g'_i(x)$ is surjective for all $i$, we can fix it as 
follows. For each $i$, we fix a lattice $\tilde H_i \subset H$ such that 
Lemma \ref{lem:main} applies to $g_i$, $x$ and $\tilde H_i$. Under our 
assumption, such lattices always exist and can be computed dynamically 
using Proposition \ref{prop:locanalytic}. Now, the equality $g_i(x + 
\tilde H_i) = x_i + g'_i(x)(\tilde H_i)$ means that any perturbation of 
$x_i$ by an element in $g'_i(x) (\tilde H_i)$ is induced by a 
perturbation of $x$ by an element in $\tilde H_i \subset H$. Hence, we 
can freely compute $x_i$ modulo $g'_i(x) (\tilde H_i)$ without changing 
the final result. Since $g'_i(x)(\tilde H_i)$ is a lattice in $E_i$, this 
remark makes possible the computation of $x_i$.

@@

For $i \in \{1, \ldots, n\}$, define $h_i = f_n \circ \cdots \circ 
f_{i+1}$, so that we have $f = h_i \circ g_i$. On differentials, we 
have $f'(x) = h_i'(x_i) \circ g'_i(x)$. Since $f'(x)$ is surjective (by 
assumption), we deduce that all $h'_i(x_i)$ are surjective. Let $H'_i$ 
be a lattice in $E_i$ such that:
\begin{enumerate}[(a)] 
\item \label{item:Hi1}
$H'_i$ is contained in $H_i + \ker h'_i(x_i) = h'_i(x_i)^{-1}
\big(f'(x)(H)\big)$;
\item \label{item:Hi2}
Lemma \ref{lem:main} applies with the function $h_i$, the point
$x_i$ and the lattice $H'_i$.
\end{enumerate}
Applying Lemma \ref{lem:main}, we then get:
$$h_i(x_i + H'_i) = x_n + h'_i(x_i)(H'_i) \subset x_n + f'(x)(H).$$
Therefore, if after the $i$-th step of the execution of the function
{\tt f}, we modify the intermediate value $x_i$ by an element of
$H'_i$, the final result remains unchanged. In other words, it is
enough to compute $x_i$ modulo $H'_i$.

@@

\todofor{Xavier}{Extend Lemma \ref{lem:main} to manifolds.}

>>>>

@@

\subsection{The case of locally analytic functions}
\label{ssec:locanalytic}

@@

Let us assume now that $f$ is locally analytic on $U$ and $H$ is a first 
order lattice for $f$ at $x$. Then $f(x+H) = f(x) + f'(x)(H)$ and the 
smallest possible $H'$ is $f'(x)(H)$. In other words, the function {\tt 
f} generates artificial loss of precision if the precision $H'$ it 
outputs is strictly larger that $f'(x)(H)$. The aim of this section is 
to explain how, under the above hypothesis, one can implement the 
function {\tt f} so that it always outputs the optimal precision $O(H')$ 
with $H' = f'(x)(H)$.

@@

There is nevertheless an important issue with this approach: in order to 
be sure that Lemma \ref{lem:main} applies, we need \emph{a priori} to 
compute the exact values of all $x_i$'s, which is of course not 
possible! Assuming that $g'_i(x)$ is surjective for all $i$, we can fix 
it as follows. For each $i$, we fix a first order lattice $\tilde H_i$ 
for $g_i$ at $x$. Under our assumption, such lattices always exist and 
can be computed dynamically using Proposition \ref{prop:locanalytic}. 
Now, the equality $g_i(x + \tilde H_i) = x_i + g'_i(x)(\tilde H_i)$ 
means that any perturbation of $x_i$ by an element in $g'_i(x) (\tilde 
H_i)$ is induced by a perturbation of $x$ by an element in $\tilde H_i 
\subset H$. Hence, we can freely compute $x_i$ modulo $g'_i(x) (\tilde 
H_i)$ without changing the final result. Since $g'_i(x)(\tilde H_i)$ is 
a lattice in $E_i$, this remark makes possible the computation of $x_i$.

@@

For $i \in \{1, \ldots, n\}$, define $h_i = f_n \circ \cdots \circ 
f_{i+1}$, so that we have $f = h_i \circ g_i$. On differentials, we 
have $f'(x) = h_i'(x_i) \circ g'_i(x)$. Since $f'(x)$ is surjective (by 
assumption), we deduce that all $h'_i(x_i)$ are surjective. Let $H'_i$ 
be a lattice in $E_i$ such that:
\begin{enumerate}[(a)] 
\item \label{item:Hi1}
$H'_i$ is contained in $H_i + \ker h'_i(x_i) = h'_i(x_i)^{-1}
\big(f'(x)(H)\big)$;
\item \label{item:Hi2}
$H'_i$ is a first order lattice for $h_i$ at $x_i$.
\end{enumerate}
By definition, we have
$h_i(x_i + H'_i) = x_n + h'_i(x_i)(H'_i) \subset x_n + f'(x)(H)$.
Therefore, if after the $i$-th step of the execution of the function
{\tt f}, we modify the intermediate value $x_i$ by an element of
$H'_i$, the final result remains unchanged. In other words, it is
enough to compute $x_i$ modulo $H'_i$.

@@

With above definitions, Lemma \ref{lem:main} extends readily to 
manifolds.

\begin{lem}
Let $V$ and $W$ be two $K$-manifolds. 
Suppose that we are given a differentiable function $f : V \to W$, 
together with a point $x \in V$ such that $f'(x) : T_x V \to T_{f(x)} W$ 
is surjective. Let $H$ be a precision datum at $x$. Then, there exists
a positive real number $\delta$ such that, for all $\lambda \in 
B_K(\delta)$, the equality
$$f(x + \lambda H) = f(x) + \lambda \cdot f'(x)(H)$$
holds.
\end{lem}

\begin{proof}
Apply Lemma \ref{lem:main} in charts.
\end{proof}

\begin{rmk}
If $V$ and $W$ are locally analytic $K$-manifolds and the mapping $f$ is 
locally analytic as well, then one can explicitely determine, using the 
method of \S \ref{ssec:locanalytic}, a value $\delta$ for which the 
above Lemma works.
\end{rmk}
